Abstractions and exemplars: the measure noun phrase alternation in German

### 2018-01-11 Latex/article.tex
About to submit revision.
Fixing replies to comments.
Minor corrections.

\runninglinenumbers
\modulolinenumbers[1]
\setlength\linenumbersep{20pt}

@@

  \abstract{In this paper, an alternation in German measure noun phrases is examined under a varying-abstraction perspective.
    In a specific measure NP construction, the embedded kind-denoting noun either agrees in case with the measure noun (\textit{eine Tasse guter Kaffee} `a cup of good coffee') or it stands in the genitive (\textit{eine Tasse guten Kaffees}).
    Each of the two alternants is syntactically similar to a non-alternating variant.
    I propose a prototype model which assigns a common prototypical meaning to each of the alternants and its corresponding non-alternating construction.
    Based on that, I predict that lexical, morpho-syntactic, and stylistic features help to predict the choice of the alternant.
    A large corpus study is presented which supports this theory.
    However, in addition to the prototype effects, an exemplar effect is also shown to influence the choice, namely the relative frequencies with which lemmas occur in the non-alternating constructions.
    I argue that allowing both prototype and exemplar effects is more adequate than following radical prototype or exemplar approaches.
    It is also verified in two experiments that the corpus-derived model corresponds to the behaviour of native speakers.
    The weak effect size of the experimental validation is discussed in the context of corpus-based cognitive linguistics and the validation of corpus-derived models.
  }
  \keywords{prototypes vs.\ exemplars, corpus methods and experimental validation, alternations, hierarchical models, pseudo-partitives, German}
  %\classification[PACS]{...}
  %\communicated{...}
  %\dedication{...}
  %\received{...}
  %\accepted{...}
  %\journalname{Submitted April 2017, revised version submitted January 2018. Do not quote.}
  %\journalyear{}
  %\journalvolume{...}
  %\journalissue{...}
  %\startpage{...}
  %\aop
  %\DOI{...}

>>>>

% \runninglinenumbers
% \modulolinenumbers[1]
% \setlength\linenumbersep{20pt}

@@

  \abstract{In this paper, an alternation in German measure noun phrases is examined under a varying-abstraction perspective.
    In a specific measure NP construction, the embedded kind-denoting noun either agrees in case with the measure noun (\textit{eine Tasse guter Kaffee} `a cup of good coffee') or it stands in the genitive (\textit{eine Tasse guten Kaffees}).
    Each of the two alternants is syntactically similar to a non-alternating construction.
    I propose a prototype model which assigns a common prototypical meaning to each of the alternants and its corresponding non-alternating construction.
    Based on this, I argue that lexical, morpho-syntactic, and stylistic features help to predict the choice of the alternant.
    A large corpus study is presented which supports this theory.
    However, in addition to the prototype effects, an exemplar effect is also shown to influence the choice, namely the relative frequencies with which lemmas occur in the non-alternating constructions.
    I argue that allowing both prototype and exemplar effects is more adequate than following radical prototype or exemplar approaches.
    It is also verified in two experiments that the corpus-derived model corresponds to the behaviour of native speakers.
    The weak effect size of the experimental validation is discussed in the context of corpus-based cognitive linguistics and the validation of corpus-derived models.
  }
  \keywords{prototypes vs.\ exemplars, corpus methods and experimental validation, alternations, hierarchical models, pseudo-partitives, German}
  %\classification[PACS]{...}
  %\communicated{...}
  %\dedication{...}
  %\received{...}
  %\accepted{...}
  %\journalname{Submitted April 2017, revised version submitted January 2018. Do not quote.}
  %\journalyear{}
  %\journalvolume{...}
  %\journalissue{...}
  %\startpage{...}
  %\aop
  %\DOI{...}

### 2018-01-11 Latex/sections/1intro.tex
Minor corrections.

On the other hand, the trade-off one has to accept when doing experiments with highly simplified stimuli and very simple tasks is their lower \textit{external validity} (\ie\ their lower degree of generalisability) and their high dependence on potentially problematic operationalisations of constructs, control of confounding factors in the face of a limited number of available subjects, etc. (in other words, critical dependence on \textit{construct validity} and \textit{internal validity}).%
\footnote{An accessible overview of the different types of validity can be found in Chapter~1 of \citet{MaxwellDelaney2004}.}
Tasks in cognitive science have been criticised exactly for their lack of external validity, for example by \citet{Murphy2003}.
From a linguistic perspective, it is remarkable in this context that \cite{VoorspoelsEa2011} consider their experimental task -- which is the assignment of typicality scores to nouns from the domains of \textit{animals} and \textit{artefacts} to categories like \textit{bird}, \textit{fish}, \textit{clothing}, or \textit{tools} -- a study of ``superordinate natural language categories, whereas most evidence supporting exemplar representations has been found in artificial categories of a more subordinate level'' \citep[1013]{VoorspoelsEa2011}.
Corpus linguists interested in probabilistic alternation modelling deal with much more complex high-level categories and use large and complex feature sets, especially in (morpho-)syntax.%
\footnote{Notice, however, that recently, approaches have emerged which solve at least some problems by abandoning linguistic high-level features altogether \citep{BaayenEa2016,RamscarPort2016}.
Clearly, they have not (or at least not yet) reached mainstream popularity, and it remains to be seen how well they perform on a broader range of questions.}
It is thus an advantage of much linguistic work on categorisation that it deals with complex and realistically produced data, because this greatly improves the external validity of studies, although by sacrificing some construct validity.
An ideal contribution by cognitive corpus linguists to the research on (levels of) category abstraction in the human mind would thus be to provide analyses which have great external validity and complexity while carefully making sure that (and determining to what extent) these finding correlate with reactions from cognitive agents under more controlled experimental conditions, which increases the construct validity.

>>>>

On the other hand, the trade-off one has to accept when doing experiments with highly simplified stimuli and very simple tasks is their lower \textit{external validity} (\ie\ their lower degree of generalisability) and their high dependence on potentially problematic operationalisations of constructs, control of confounding factors in the face of a limited number of available subjects, etc. (in other words, critical dependence on \textit{construct validity} and \textit{internal validity}).%
\footnote{An accessible overview of the different types of validity can be found in Chapter~1 of \citet{MaxwellDelaney2004}.}
Tasks in cognitive science have been criticised exactly for their lack of external validity, for example by \citet{Murphy2003}.
From a linguistic perspective, it is remarkable in this context that \cite{VoorspoelsEa2011} consider their experimental task -- which is the assignment of typicality scores to nouns from the domains of \textit{animals} and \textit{artefacts} to categories like \textit{bird}, \textit{fish}, \textit{clothing}, or \textit{tools} -- a study of ``superordinate natural language categories, whereas most evidence supporting exemplar representations has been found in artificial categories of a more subordinate level'' \citep[1013]{VoorspoelsEa2011}.
Corpus linguists interested in probabilistic alternation modelling deal with much more complex high-level categories and use large and complex feature sets, especially in (morpho-)syntax.%
\footnote{Notice, however, that recently, approaches have emerged which solve at least some problems by abandoning linguistic high-level features altogether \citep{BaayenEa2016,RamscarPort2016}.
Clearly, they have not (or at least not yet) reached mainstream popularity, and it remains to be seen how well they perform on a broader range of questions.}
It is thus an advantage of much linguistic work on categorisation that it deals with complex and realistically produced data because this greatly improves the external validity of studies, although by sacrificing some construct validity.
An ideal contribution by cognitive corpus linguists to the research on (levels of) category abstraction in the human mind would thus be to provide analyses which have great external validity and complexity while carefully making sure that (and determining to what extent) these finding correlate with reactions from cognitive agents under more controlled experimental conditions, which increases the construct validity.
